---
title: "Wine Quality EDA - Capstone"
author: "Kasey Cohen"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Business Understanding

A wine producer wants to understand what factors influence wine quality. This analysis explores non-obvious relationships between chemical properties of wine and its quality using EDA techniques.

# 2. Data Understanding

```{r}
# Load libraries
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("ggcorrplot")) install.packages("ggcorrplot")
if (!require("skimr")) install.packages("skimr")
if (!require("ggpubr")) install.packages("ggpubr")

library(tidyverse)
library(ggcorrplot)
library(skimr)
library(ggpubr)

# Load dataset
wine <- read.csv("https://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-red.csv", sep = ";")
head(wine)
```

```{r}
# Basic structure
dim(wine)
str(wine)
skim(wine)
summary(wine)
```

# 3. Data Preparation

```{r}
# Check for missing values
colSums(is.na(wine))

# Create quality labels
wine <- wine %>%
  mutate(quality_label = case_when(
    quality <= 5 ~ "low",
    quality == 6 ~ "medium",
    quality >= 7 ~ "high"
  )) %>%
  mutate(quality_label = factor(quality_label, levels = c("low", "medium", "high")))

table(wine$quality_label)
```

# 4. Modeling and Evaluation (EDA)

## Correlation Heatmap

```{r}
cor_matrix <- cor(wine[, 1:11])
ggcorrplot(cor_matrix, hc.order = TRUE, type = "lower", lab = TRUE)
```

## Boxplot: Alcohol vs Quality

```{r}
ggplot(wine, aes(x = quality_label, y = alcohol, fill = quality_label)) +
  geom_boxplot() +
  labs(title = "Alcohol Content by Wine Quality", x = "Quality", y = "Alcohol") +
  theme_minimal()
```

## Density Plot: pH

```{r}
ggplot(wine, aes(x = pH, fill = quality_label)) +
  geom_density(alpha = 0.6) +
  labs(title = "pH Distribution by Wine Quality", x = "pH", y = "Density") +
  theme_minimal()
```

## Hypothesis Testing

```{r}
# T-test: Alcohol in high vs. low quality
t.test(alcohol ~ quality_label, data = wine %>% filter(quality_label != "medium"))

# ANOVA: Fixed acidity vs. quality
anova_result <- aov(fixed.acidity ~ quality_label, data = wine)
summary(anova_result)

# QQ plot
ggqqplot(wine$alcohol)
```

# 5. Conclusions

```{r}
cat("Key Insights:\n")
cat("- Alcohol is positively correlated with wine quality.\n")
cat("- Volatile acidity is inversely related to quality.\n")
cat("- Minor relationships observed with pH and citric acid.\n")
```
